---
title: "Writing functions in R"
output: html_document
---

## Your Turn 0

In the setup chunk below, load the `tidyverse` packages.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

When doing your data analysis, you may find yourself copying and pasting your code to replicate your analysis on a new version of the dataset or on a new variable. In those cases, the ideal way to proceed is to covert your code into a function, then reuse that function to conduct the analsis of a different dataset. This will save you a lot of copying and pasting, and help troubleshoot the problems when they arise.

## Your Turn 1

- The code chunk below contains the function `fahrenheit_to_celsius`. Modify this code to create a function called `celsius_to_fahrenheit` that will convert the temperature from celius to fahrenheit. The formula to convert the temperatre from celsius to fahrenheit is `temp_F = (temp_C * 9/5) + 32`.

- Using your newly created `celsius_to_fahrenheit` function, convert 0 C to fahrenheit. Your result should equal to 32.

```{r}
# a function to convert temperature from fahrenheit to celsius
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
  return(temp_C)
}

```


## Your Turn 2

Load the package **readxl**, and use the function `read_excel()` to import `PLPTH850dataset_jamesLin.xlsx` into R, and name it `soil`. This is a dataset shared by one of your class mate, James. We will use this dataset to demonstrate writing and using functions.

```{r}
library(readxl)
soil <- read_excel("PLPTH850dataset_jamesLin.xlsx")
```

Confirm that `soil` dataset has 208 observations and 39 variables. You can look into Environment tab on the top right or use the function `str()` or `glimpse()` on the dataset `soil`.

## Your Turn 3

Some of the variable names in `soil` contain special characters such `%, +, -,(, )`. Lets use a package called **janitor** to clean variable (column) names. In the code chunk below:

- Load the package **janitor**

- Use the funcion `clean_names()` to clean column names of `soil` datset

```{r}
library(janitor)

soil <- soil %>% 
  clean_names()
```

## Your Turn 4

**Write a simple function**

Here is a simple syntax for creating a function:

```{r}
function_name <- function(argument1, argument2, ...){
 # function code goes here
 # here, you will write what you would like your function to do
}
```

To use the function, you will run `function_name(argument1, argument2)`. 

Lets build a simple function, that takes in a variable name from `soil` and computes its mean. Name this function `calculate_mean`, and it will have only one argument. **Hint**: you can modify the function written below to calculate the sum of a given vairable in `soil` dataset.

```{r}
calculate_sum <- function(variable){
  sum(soil[[variable]], na.rm = TRUE)
}

calculate_sum("ca")

```


## Your Turn 5

To take advantage of writing functions for tidyverse code, you will need to load **rlang** package to write functions in tidyverse format. So, in the code chunk below:

- Load the package **rlang**.

- One trick to write a function is to first write the code for an actual variable, and then replace that variable name with a function argument. I will write a function below to calculate the sum of a variable grouped by treatment and depth. Later you will modify it to calculate mean of a given variable by grouped by treatment and depth.

```{r}
# Load **rlang**
library(rlang)

# test what you want to do on an actual variable
soil %>% 
  group_by(treatment, depth) %>% 
  summarise(mean_ca = mean(ca, na.rm = TRUE))

# the above code worked, it gave sum for each combination
# of treatment and depth
# Now, I would like to put it in a function

calculate_sum <- function(){
  # function code goes here
}

# copy the code from line 99 to 101 and paste it in the 
# function skeleton below, and replace `ca` with
# `variable`

calculate_sum <- function(variable){
  soil %>% 
  group_by(treatment, depth) %>% 
  summarise(sum_variable = sum(variable, na.rm = TRUE))
}

calculate_sum(ca)

# you will encounter the following error
# Error: object 'ca' not found

# Let's fix it: put `variable` inside double curly brackets

calculate_sum <- function(variable){
  soil %>%
  group_by(treatment, depth) %>%
  summarise("mean_{{variable}}" := mean({{variable}}, na.rm = TRUE))
}

calculate_sum(ca)

# you can also provide the variable name by naming the argument
calculate_sum(variable = toc_g_kg)
```

In the chunk below, modify the the `calculate_sum()` function written in the above chunk to create `calculate_mean()`.

```{r}

```



