---
title: "Tidy Data"
---

```{r setup}
library(tidyverse)
library(babynames)

# Toy data
cases <- tribble(
  ~Country, ~"2011", ~"2012", ~"2013",
      "FR",    7000,    6900,    7000,
      "DE",    5800,    6000,    6200,
      "US",   15000,   14000,   13000
)

pollution <- tribble(
       ~city, ~size, ~amount,
  "New York", "large",      23,
  "New York", "small",      14,
    "London", "large",      22,
    "London", "small",      16,
   "Beijing", "large",     121,
   "Beijing", "small",     121
)

names(who) <- stringr::str_replace(names(who), 
                                   "newrel", 
                                   "new_rel")
```

```{r}
table1
table2
table3
table4a
table4b
table5
```

# tidyr


## Your Turn 1

On a sheet of paper, draw how the cases data set would look if it had the same values grouped into three columns: **country**, **year**, **n**

## Your Turn 2

Use `pivot_longer()` to reorganize `table4a` into three columns: **country**, **year**, and **cases**.

```{r}
table4a %>% 
  pivot_longer(cols = `1999`:`2000`,
               names_to = "year",
               values_to = "cases")
```

## Your Turn 3

On a sheet of paper, draw how this data set would look if it had the same values grouped into three columns: **city**, **large**, **small**

## Your Turn 4

Use `pivot_wider()` to reorganize `table2` into four columns: **country**, **year**, **cases**, and **population**.

```{r}
table2 %>% 
  pivot_wider(names_from = type,
              values_from = count)

```

# Practice on a real agricultural dataset

```{r}
soil <- read_excel("PLPTH850dataset_jamesLin.xlsx") %>% 
  View()
```

## Your Turn 5

- Select columns `SampleName` through `replication`, and `Ca` through `Fe`.

- Then use `pivot_longer()` to put column names of `Ca:Fe` into a column named `Nutrient` and the corresponding values/data into a column named `conc_ppm`.

- Name the resulting dataset `soil_long`

```{r}
soil_long <- soil %>% 
  select(SampleName:replication, Ca:Fe) %>% 
  pivot_longer(cols = Ca:Fe, names_to = "Nutrient", values_to = "conc_ppm")

soil_long
```


## Your Turn 6

In the `soil` dataset, separate the `Depth` column in to `depth_low` and `depth_high`.

*(Hint: Be sure to remove each `_` before running the code)*

```{r}
soil %>%
  separate(_____, into = c(_______, _______), sep = _____) %>%
```


***

# Take Aways

Data comes in many formats but R prefers just one: _tidy data_.

A data set is tidy if and only if:

1. Every variable is in its own column
2. Every observation is in its own row
3. Every value is in its own cell (which follows from the above)

What is a variable and an observation may depend on your immediate goal.

